Generalized spaces for constructive algebra

### 2020-04-08 paper-herrsching2019/main.tex
Two more XXX's

\begin{abstract}
XXX
\end{abstract}
\body

@@

The completeness theorem for coherent logic~\cite[Corollary~D1.5.10]{johnstone:elephant} even
gives an a priori reason why (for a given matrix size) a finitary proof has to
exist. However, since the proof of the completeness theorem itself uses~\BPIT
(and indeed is equivalent to it), it does not give any indication how such a
proof could be found. (Still, beautiful constructive proofs are known and
presented, for instance, in a celebrated short note by Fred Richman on nontrivial
uses of trivial rings~\cite{richman:trivial-rings} and in the recent textbook by
Henri Lombardi and Claude Quitté on constructive commutative
algebra~\cite{lombardi-quitte:constructive-algebra}.)
% FUTURE: Find precise reference.

@@

In some cases, these new techniques XXX

@@

\paragraph{Extracting constructive content from classical proofs.}\label{par:barr}
Geometric logic strikes a fine balance between being sufficiently expressive in order to
have rich applications and also being sufficiently restrictive so that \emph{Barr's
theorem} holds. This theorem is concerned with eliminating classical reasoning
from proofs of geometric sequents, and it comes in a simple yet still useful
and a more sophisticated version:
\begin{enumerate}
\item If infinitary classical logic proves a geometric sequent, then so does
geometric logic.
\item If a specific version of extensional type theory with the law of excluded
middle and the axiom of choice proves a geometric sequent, then so does
geometric logic.
\end{enumerate}
For instance, the \emph{Nullstellensatz} is the statement that for any
system~$f_1,\ldots,f_n \in K[X_1,\ldots,X_m]$ of polynomials over an
algebraically closed field~$K$, either the polynomials~$f_i$ have a common zero
or there are polynomials~$g_i$ such that~$1 = \sum_i f_i g_i$ (such an
expression is an \emph{algebraic witness} of the nonexistence of a common
zero).

>>>>

\begin{abstract}
XXX abstract
\end{abstract}
\body

@@

The completeness theorem for coherent logic~\cite[Corollary~D1.5.10]{johnstone:elephant} even
gives an a priori reason why (for a given matrix size) a finitary proof has to
exist. However, since the proof of the completeness theorem itself uses~\BPIT
(and indeed is equivalent to it), it does not give any indication how such a
proof could be found. (Still, beautiful constructive proofs are known and
presented, for instance, in a celebrated short note by Fred Richman on nontrivial
uses of trivial rings~\cite{richman:trivial-rings} and in the recent
textbook~\cite{lombardi-quitte:constructive-algebra} by Henri Lombardi and
Claude Quitté on constructive commutative algebra.)
% FUTURE: Find precise reference.

@@

In some cases, these new techniques go even beyond their classical role models.
For instance, while a module need not be free ``on a dense open'' if all its
stalks at minimal prime ideals are, a module~$M$ is free in that sense if its
mirror image~$M^\sim$ is \notnot free. This observation is the basis for a
new proof of Grothendieck's generic freeness lemma, reviewed in
Section~\ref{sect:example-applications}, which is shorter and arguably more
perspicious than all previously known classical proofs.

@@

\paragraph{Extracting constructive content from classical proofs.}\label{par:barr}
Geometric logic strikes a fine balance between being sufficiently expressive in order to
have rich applications and also being sufficiently restrictive so that \emph{Barr's
theorem} holds. This theorem is concerned with eliminating classical reasoning
from proofs of geometric sequents, and it comes in a simple yet still useful
and a more sophisticated version:
\begin{enumerate}
\item If infinitary classical logic proves a geometric sequent, then so does
geometric logic.
\item If a specific version of extensional type theory with the law of excluded
middle and the axiom of choice proves a geometric sequent, then so does
geometric logic.
\end{enumerate}
Barr's theorem has ramifications to constructive algebra, as pointed out by
Gavin Wraith~\cite{wraith:intuitionistic-algebra}.
For instance, the \emph{Nullstellensatz} is the statement that for any
system~$f_1,\ldots,f_n \in K[X_1,\ldots,X_m]$ of polynomials over an
algebraically closed field~$K$, either the polynomials~$f_i$ have a common zero
or there are polynomials~$g_i$ such that~$1 = \sum_i f_i g_i$ (such an
expression is an \emph{algebraic witness} of the nonexistence of a common
zero).
